---
title: "Ejemplos Ecuaciones Diferenciales Ordinarias"
author: "Análisis Numérico"
date: "Abril 2024"
output: 
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---





<style>
body {
text-align: justify}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(pracma)
library(plotly)
library(polynom)
```



# Ejemplo 1

Consideremos la siguiente función

 $f(x)=x\, sen\,x+x^2cos\,x$, $x\in [0,\pi]$


La derivada
$f'(x)=3xcos(x)+sin(x)-x^2sin(x)$



# Ejemplo 2

Consideramos el problema de valor inicial

\begin{equation}
\frac{dy}{dt}=y-t^2+1, \qquad 0\leq t \leq 2, \qquad y(0)=0.5
\end{equation}


```{r}
dy <- function(t,y){
  y-t^2+1
}
y_sol <- function(t){
  -exp(t)/2+(t+1)^2
}

my_euler <- function(dy, a,b, N, y0){
  h <- (b-a)/N
  T <- a
  W <- y0
  for (k in 1:N) {
    w_new <- W[k]+ h*dy(T[k], W[k])
    T <- c(T, T[k]+h)
    W <- c(W, w_new)
  }
  resultados <- list("T"=T, "W"=W)
  return(resultados)
}

# h <- 0.5 
# T <- 0
# W <- 0.5
# t1 <- 0+h
# w1 <- 0.5+h*dy(0, 0.5)
# T <- c(T, t1)
# W <- c(W,w1)
# 
# t2 <- t1+h
# w2 <- w1+h*dy(t1, w1)
# T <- c(T, t2)
# W <- c(W,w2)
# 
# t3 <- t2+h
# w3 <- w2+h*dy(t2, w2)
# T <- c(T, t3)
# W <- c(W,w3)
# 
# 
# t4 <- t3+h
# w4 <- w3+h*dy(t3, w3)
# T <- c(T, t4)
# W <- c(W,w4)

sol_aprox <- my_euler(dy, 0, 2, 100, 0.5)


my_colors=c("y(x)"="firebrick", "w(x)"="dodgerblue")

ggplot()+
  geom_function(aes(color="y(x)"),fun=y_sol, xlim=c(0,2))+
  geom_line(aes(sol_aprox$T,sol_aprox$W, color="w(x)"))+
  #geom_point(aes(sol_aprox$T,sol_aprox$W), color="green", size=3)+
  scale_color_manual(values=my_colors)+
  labs(x="t", y="y(t)", color=" ")+
  theme_bw()

```

```{r}
f <- function(t,y){
  y-t^2+1
}
df <- function(t,y){
  -2*t+f(t,y)
}

ddf <- function(t,y){
  4*t+f(t,y)
}


my_taylor2 <- function(f, df, ddf , a,b, N, y0){
  h <- (b-a)/N
  T <- a
  W <- y0
  for (k in 1:N) {
    w_new <- W[k]+ h*f(T[k], W[k])+h^2/2*df(T[k], W[k])+h^3/6*ddf(T[k], W[k])
    T <- c(T, T[k]+h)
    W <- c(W, w_new)
  }
  resultados <- list("T"=T, "W"=W)
  return(resultados)
}



sol_aprox_2 <- my_taylor2(f, df, ddf, 0, 2, 100, 0.5)


my_colors=c("y(x)"="firebrick", "w(x)"="dodgerblue", "w_2(x)"="deeppink3")

sol_graf <- ggplot()+
  geom_function(aes(color="y(x)"),fun=y_sol, xlim=c(0,2))+
  geom_line(aes(sol_aprox$T,sol_aprox$W, color="w(x)"))+
  geom_line(aes(sol_aprox_2$T,sol_aprox_2$W, color="w_2(x)"))+
  #geom_point(aes(sol_aprox$T,sol_aprox$W), color="green", size=3)+
  scale_color_manual(values=my_colors)+
  labs(x="t", y="y(t)", color=" ")+
  theme_bw()


ggplotly(sol_graf)
```



## Gráficas del error y de la cota


```{r}
error_real <- y_sol(sol_aprox$T)-sol_aprox$W


M <- abs(-exp(2)/2+2)
L <- 1
h <- 0.02
cota <- h*M/(2*L)*(exp(L*(sol_aprox$T))-1)



graf_error <- ggplot()+
  geom_line(aes(sol_aprox$T, error_real), color="gold")+
  geom_point(aes(sol_aprox$T, error_real), color="red2")+
  geom_line(aes(sol_aprox$T, cota), color="blue")+
  theme_bw()

ggplotly(graf_error)
```




# Ejemplo 3


Considerermos el PVI:

\begin{equation}
\frac{dy}{dy}=te^{3t}-2y, \qquad 0\leq t \leq 2, \qquad y(0)=0 
\end{equation}


```{r}
y_sol <- function(t){
  1/25*exp(-2*t)+1/5*t*exp(3*t)-1/25*exp(3*t)
}
f <- function(t,y){
  t*exp(3*t)-2*y
}


sol_aprox <- euler_heun(f, 0, 2, 0,n=10, improved = FALSE)

my_colors=c("y(x)"="firebrick", "w(x)"="dodgerblue")

ggplot()+
  geom_function(aes(color="y(x)"),fun=y_sol, xlim=c(0,2))+
  geom_line(aes(sol_aprox$t,sol_aprox$y, color="w(x)"))+
  #geom_point(aes(sol_aprox$T,sol_aprox$W), color="green", size=3)+
  scale_color_manual(values=my_colors)+
  labs(x="t", y="y(t)", color=" ")+
  theme_bw()

```

## Gráficas del error y de la cota


```{r}
error_real <- y_sol(sol_aprox$t)-sol_aprox$y


M <- 3*exp(6)
L <- 2
h <- 0.2
cota <- h*M/(2*L)*(exp(L*(sol_aprox$t))-1)



graf_error <- ggplot()+
  geom_line(aes(sol_aprox$t, error_real), color="gold")+
  geom_point(aes(sol_aprox$t, error_real), color="red")+
  geom_line(aes(sol_aprox$t, cota), color="blue")+
  theme_bw()

ggplotly(graf_error)
```



Ejercicio 3

Suppose that we use the Improved Euler's method to approximate the solution to the differential equation



$ \frac{dy}{dx}=x−0.5y;\quad y(0.2)=4$


Let $f(x,y)=x−0.5y.$ We let $x_0=0.2$ and $y_0=4$ and pick a step size h=0.25.
 The improved Euler method is the the following algorithm. From $(x_n,y_n)$,
 our approximation to the solution of the differential equation at the n
-th stage, we find the next stage by computing the x
-step $xn+1=xn+h$,
 and then $k1$,
 the slope at $(xn,yn)$.
 The predicted new value of the solution is zn+1=yn+h⋅k1.
 Then we find the slope at the predicted new point k2=f(xn+1,zn+1)
 and get the corrected point by averaging slopes
yn+1=yn+h2(k1+k2).
Complete the following table:


```{r}
f <- function(x,y){
  x-0.5*y
}
y_sol <- function(x){
  7.6*exp(0.1)*exp(-x/2)+2*x-4
}
  
  

euler_modificado <- function(f, a, b, N, y0){
  h <- (b-a)/N
  X <- seq(a, b, by=h)
  Y <- y0
  K1 <- numeric()
  K2 <- numeric()
  Z <- numeric()
  for (k in 1:N) {
    K1[k] <- f(X[k],Y[k])
    Z[k]=Y[k]+h*K1[k]
    K2[k]=f(X[k+1], Z[k])
     Y[k+1] <- Y[k]+h/2*(K1[k]+ K2[k]) 
    # Y[k+1] <- Y[k]+h/2*(f(X[k],Y[k])+ f(X[k+1], Y[k]+h*f(X[k],Y[k]))) 
  }
  
  resultados <- list("X"=X, "Y"=Y, "K1"=K1, "Z"=Z,  "K2"=K2)
  return(resultados)
  
}

sol_aprox <- euler_modificado(f, 0.2, 1.2, 4,4)


graf_ej3 <- ggplot()+
  geom_function(fun=y_sol, xlim=c(0.2, 1.2), color="red4", linewidth=1.2)+
  geom_line(aes(sol_aprox$X, sol_aprox$Y), color="dodgerblue", linewith=1)+
  theme_test()

ggplotly(graf_ej3)

```
























