---
title: "Series de Tiempo. Indicadores."
author: "Análisis Numérico"
date: "Marzo 2024"
output: 
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---





<style>
body {
text-align: justify}
</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(lubridate)
library(pracma)
library(plotly)
library(MASS)
library(ISLR2)
library(car)
library(corrplot)
library(rgl)
library(readxl)

library(xts)
library(TSstudio)
knitr::knit_hooks$set(webgl = hook_webgl)
```



```{r}
datos <- read_xlsx("Indicadores.xlsx")
datos$Periodo <- ym(datos$Periodo)
#datos <- datos[ , 1:7]
glimpse(datos)
```


# Cetes 28 días {.tabset}

## Distribución 

```{r echo=FALSE}
data_Cetes_28_dias <- datos |> dplyr::select(Periodo, Cetes_28_dias)
data_Cetes_28_dias <- na.omit(data_Cetes_28_dias)

ggplot(data_Cetes_28_dias)+
  geom_histogram(aes(data_Cetes_28_dias$Cetes_28_dias, fill=factor(year(Periodo))), color="black")+
  labs(fill="Año")+
  theme_bw()
```




## Serie temporal

```{r echo=FALSE}
Cetes_28_dias_ts <- ts(data_Cetes_28_dias$Cetes_28_dias, start = c(1993, 1), frequency = 12)
Cetes_28_dias_xts <- as.xts(Cetes_28_dias_ts)

ts_info(Cetes_28_dias_xts)
ts_plot(Cetes_28_dias_xts, slider = TRUE)
```

## Descomposición

```{r echo=FALSE}
Cetes_28_dias_dec <- decompose(Cetes_28_dias_ts, type="additive" )
plot(Cetes_28_dias_dec)
```


## ACF

```{r echo=FALSE}
Cetes_28_dias_acf <- acf(Cetes_28_dias_ts)
```


## Autorregresión {.tabset}

### Coeficientes

```{r echo=FALSE}
Cetes_28_dias_ar <- ar(Cetes_28_dias_ts, method = "mle", order.max = 1)
Cetes_28_dias_ar
```

### Gráfica

```{r echo=FALSE}
ts.plot(Cetes_28_dias_ts, Cetes_28_dias_ts-Cetes_28_dias_ar$resid, lty=1:2, col=c("blue", "red"))
```


### Residuos

```{r echo=FALSE}
ts_plot(Cetes_28_dias_ar$resid)
```


# Actividades terciarias {.tabset}




## Distribución 

```{r echo=FALSE}
data_actividades_terciarias <- datos |> dplyr::select(Periodo, actividades_terciarias)
data_actividades_terciarias <- na.omit(data_actividades_terciarias)

ggplot(data_actividades_terciarias)+
  geom_histogram(aes(data_actividades_terciarias$actividades_terciarias, fill=factor(year(Periodo))), color="black")+
  labs(fill="Año")+
  theme_bw()
```




## Serie temporal

```{r echo=FALSE}
actividades_terciarias_ts <- ts(data_actividades_terciarias$actividades_terciarias, start = c(1993, 1), frequency = 12)
actividades_terciarias_xts <- as.xts(actividades_terciarias_ts)

ts_info(actividades_terciarias_xts)
ts_plot(actividades_terciarias_xts, slider = TRUE)
```

## Descomposición

```{r echo=FALSE}
actividades_terciarias_dec <- decompose(actividades_terciarias_ts, type="additive" )
plot(actividades_terciarias_dec)
```


## ACF

```{r echo=FALSE}
actividades_terciarias_acf <- acf(actividades_terciarias_ts)
```


## Autorregresión {.tabset}

### Coeficientes

```{r echo=FALSE}
actividades_terciarias_ar <- ar(actividades_terciarias_ts, method = "mle", order.max = 1)
actividades_terciarias_ar
```

### Gráfica

```{r echo=FALSE}
ts.plot(actividades_terciarias_ts, actividades_terciarias_ts-actividades_terciarias_ar$resid, lty=1:2, col=c("blue", "red"))
```


### Residuos

```{r echo=FALSE}
ts_plot(actividades_terciarias_ar$resid)
```


# Tipo de cambio real {.tabset}


## Distribución 

```{r echo=FALSE}
data_tipo_cambio_real <- datos |> dplyr::select(Periodo, tipo_cambio_real)
data_tipo_cambio_real <- na.omit(data_tipo_cambio_real)

ggplot(data_tipo_cambio_real)+
  geom_histogram(aes(data_tipo_cambio_real$tipo_cambio_real, fill=factor(year(Periodo))), color="black")+
  labs(fill="Año")+
  theme_bw()
```




## Serie temporal

```{r echo=FALSE}
tipo_cambio_real_ts <- ts(data_tipo_cambio_real$tipo_cambio_real, start = c(1993, 1), frequency = 12)
tipo_cambio_real_xts <- as.xts(tipo_cambio_real_ts)

ts_info(tipo_cambio_real_xts)
ts_plot(tipo_cambio_real_xts, slider = TRUE)
```

## Descomposición

```{r echo=FALSE}
tipo_cambio_real_dec <- decompose(tipo_cambio_real_ts, type="additive" )
plot(tipo_cambio_real_dec)
```


## ACF

```{r echo=FALSE}
tipo_cambio_real_acf <- acf(tipo_cambio_real_ts)
```


## Autorregresión {.tabset}

### Coeficientes

```{r echo=FALSE}
tipo_cambio_real_ar <- ar(tipo_cambio_real_ts, method = "mle", order.max = 1)
tipo_cambio_real_ar
```

### Gráfica

```{r echo=FALSE}
ts.plot(tipo_cambio_real_ts, tipo_cambio_real_ts-tipo_cambio_real_ar$resid, lty=1:2, col=c("blue", "red"))
```


### Residuos

```{r echo=FALSE}
ts_plot(tipo_cambio_real_ar$resid)
```





# Tasa desocupación urbana {.tabset}


## Distribución 

```{r echo=FALSE}
data_tasa_desocupacion_urbana <- datos |> dplyr::select(Periodo, tasa_desocupacion_urbana)
data_tasa_desocupacion_urbana <- na.omit(data_tasa_desocupacion_urbana)

ggplot(data_tasa_desocupacion_urbana)+
  geom_histogram(aes(data_tasa_desocupacion_urbana$tasa_desocupacion_urbana, fill=factor(year(Periodo))), color="black")+
  labs(fill="Año")+
  theme_bw()
```




## Serie temporal

```{r echo=FALSE}
tasa_desocupacion_urbana_ts <- ts(data_tasa_desocupacion_urbana$tasa_desocupacion_urbana, start = c(2005, 1), frequency = 12)
tasa_desocupacion_urbana_xts <- as.xts(tasa_desocupacion_urbana_ts)

ts_info(tasa_desocupacion_urbana_xts)
ts_plot(tasa_desocupacion_urbana_xts, slider = TRUE)
```

## Descomposición

```{r echo=FALSE}
tasa_desocupacion_urbana_dec <- decompose(tasa_desocupacion_urbana_ts, type="additive" )
plot(tasa_desocupacion_urbana_dec)
```


## ACF

```{r echo=FALSE}
tasa_desocupacion_urbana_acf <- acf(tasa_desocupacion_urbana_ts)
```


## Autorregresión {.tabset}

### Coeficientes

```{r echo=FALSE}
tasa_desocupacion_urbana_ar <- ar(tasa_desocupacion_urbana_ts, method = "mle", order.max = 1)
tasa_desocupacion_urbana_ar
```

### Gráfica

```{r echo=FALSE}
ts.plot(tasa_desocupacion_urbana_ts, tasa_desocupacion_urbana_ts-tasa_desocupacion_urbana_ar$resid, lty=1:2, col=c("blue", "red"))
```


### Residuos

```{r echo=FALSE}
ts_plot(tasa_desocupacion_urbana_ar$resid)
```




# Ruido blanco {.tabset}

## Distribución 

```{r echo=FALSE}
data_white_noise <- datos |> dplyr::select(Periodo)
data_white_noise$white_noise <- rnorm(374, mean=0, sd=1)

ggplot(data_white_noise)+
  geom_histogram(aes(data_white_noise$white_noise, fill=factor(year(Periodo))), color="black")+
  labs(fill="Año")+
  theme_bw()
```




## Serie temporal

```{r echo=FALSE}
white_noise_ts <- ts(data_white_noise$white_noise, start = c(1993, 1), frequency = 12)
white_noise_xts <- as.xts(white_noise_ts)

ts_info(white_noise_xts)
ts_plot(white_noise_xts, slider = TRUE)
```

## Descomposición

```{r echo=FALSE}
white_noise_dec <- decompose(white_noise_ts, type="additive" )
plot(white_noise_dec)
```


## ACF

```{r echo=FALSE}
white_noise_acf <- acf(white_noise_ts)
```


## Autorregresión {.tabset}

### Coeficientes

```{r echo=FALSE}
white_noise_ar <- ar(white_noise_ts, method = "mle", order.max = 1)
white_noise_ar
```

### Gráfica

```{r echo=FALSE}
ts.plot(white_noise_ts, white_noise_ts-white_noise_ar$resid, lty=1:2, col=c("blue", "red"))
```


### Residuos

```{r echo=FALSE}
ts_plot(white_noise_ar$resid)
```










